Reinforcement Learning Methods for Operations Research Applications -- Reloaded

### 2020-12-18 paper/paper.tex
update literature

One way to address this nonlinear interaction effects is to set lead times dynamically. Intuitively
the order release problem is solved by perfectly matching the lead times to the flow times, but the
corresponding optimisation problem faces ``sampling issues'' meaning that the flow times depend on
the lead times. An extreme scenario of this problem is the so called ``lead time syndrome'', which
describes a vicious cycle where increasing flow times perpetually inflate the lead times which leads
to worse performance (see e.g., \citealt{Mather1978, knollmann2013control, Selcuk2006}. Therefore,
\citet{SchneckenreitherHaeussler2019} propose to use reinforcement learning as approach to
incorporate these effects in this optimisation problem. Reinforcement learning is an optimisation
technique that stems from dynamic programming and its goal is to find the best stationary policy for
a given problem. This policy is usually provided by assessing the states (or state-action pairs) of
an underlying Markov Decision Process (MDP). The advantage of reinforcement learning over dynamic
programming is that (i) the problem space is explored by an agent and thus only expectantly
interesting parts of the problem space need to be assessed and (ii) the knowledge (acquisition) of
transition probabilities becomes unnecessary as the states are evaluated by consecutively observed
states solely.

@@

Therefore, most applications that incorporate and directly reflect costs or profit in the reward
function use average reward reinforcement learning. \citet{aydin2000dynamic} use it in the setting
of scheduling, while in a series of papers Mahadevan et al. investigated several problem domains
starting with simple MDPs
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
After these foundational works for average reward reinforcement learning they introduced a
continuous time average reward reinforcement learning algorithm named SMART
\citep{Mahadevan97_SelfimprovingFactorySimulationUsingContinuoustimeAveragerewardReinforcementLearning}.
Applications of SMART reach from the optimisation of queuing systems
\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies,Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning},
maintenance of an inventory system
\citep{Das99_SolvingSemiMarkovDecisionProblemsUsingAverageRewardReinforcementLearning} to optimising
transfer line in terms of maximizing demand, while keeping inventory levels of unfinished product as
low as possible \citep{Mahadevan98_OptimizingProductionManufacturingUsingReinforcementLearning}.
However, in practise usually decision have to be made on a daily basis \citep{enns2004work}.
Therefore, we refrain from this adaption and concentrate on standard MDPs only. Furthermore, often
continuous-time semi-MDP problems can be converted through uniformisation into equivalent discrete
time instances \citep[see][]{Puterman94,bertsekas1995dynamic}.

>>>>

One way to address this nonlinear interaction effects is to set lead times dynamically. Intuitively
the order release problem is solved by perfectly matching the lead times to the flow times, but the
corresponding optimisation problem faces ``sampling issues'' meaning that the flow times depend on
the lead times. An extreme scenario of this problem is the so called ``lead time syndrome'', which
describes a vicious cycle where increasing flow times perpetually inflate the lead times which leads
to worse performance (see e.g., \citealt{Mather1978, knollmann2013control, Selcuk2006}.
%
Thus, setting lead times dynamically harbors optimisation potential \citep{hoyt1978dynamic}, but may
also substantially degrade the system performance.
%
\citet{schneckenreither2020order} have established
following categorisation of dynamic lead time management approaches:

\begin{itemize}
\item \textsf{Reactive lead time management} approaches set lead times by reacting on
  earlier flow times (e.g., \citealt{enns2004work, Selcuk2006}). Note that the forecast is always
  based on \textit{past} data as the most recent system changes cannot be reflected by flow times
  until the corresponding orders arrive in the FGI, which might take several
  periods.
\item
  \textsf{Proactive lead time management} may incorporate \textit{past} data in conjunction with
  the \textit{current} system state to set lead times (e.g., \citealt{Bertrand1983,
    ChungHuang2002}). Put differently, these methods aim to find a function that provides lead times
  based on the current state of the production system and possibly information from the past.

\item \textsf{Predictive lead time management} may not only incorporate \textit{past} data and the
  \textit{current} system state to set lead times, but also utilises the anticipated \textit{future}
  system state to detect arising issues of future periods and react accordingly (e.g.
  \citealt{PaterninaArboleda2001,schneckenreither2020order,Schneckenreither2019}). Thus, it extends
  proactive lead time management from a flow time forecasting or simple lead time setting technique
  to a lead time management approach that integrates the future behaviour of the system when setting
  lead times. This allows reasoning of the system dynamics, as for instance triggering the lead time
  system, and thus such an algorithm can react accordingly to find a more farsighted optimal lead
  time update.
\end{itemize}

We hypothesis that dynamic lead time management approaches need to aim for a predictive design in
order to be able to compete with state-of-the-art order release methods from literature. However,
there only exist three papers that propose a predictive lead time management algorithm in
literature.

The first approach by \citet{PaterninaArboleda2001} introduces a predictive order release model by
using reinforcement learning in a single product, four-station serial flow line and compare its
performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP).
%
% Reinforcement learning is an optimisation technique that stems from dynamic programming and its goal
% is to find the best stationary policy for a given problem. This policy is usually provided by
% assessing current and future states (or state-action pairs) of an underlying Markov Decision Process
% (MDP).
% The advantage of reinforcement learning over dynamic programming is that (i) the problem
% space is explored by an agent and thus only expectantly interesting parts of the problem space need
% to be assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary
% as the states are evaluated by consecutively observed states solely.
%
The algorithm of \citet{PaterninaArboleda2001} decides on whether or not to release an order after
each system change, the completion of an operation of any order at any stage or a new order arrival,
and assumes that any unsatisfied demand is lost. Thus, they use a continuous order release method
although in practice order release decisions often need to be made on a periodical basis, e.g.\@
daily (see \citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their tabular
based reinforcement learning agent.
%
Then, \citet{Schneckenreither2019} use several different reinforcement learning algorithms to make
periodic order release decisions for a flow shop production system. The algorithm directly sets lead
times for each product type, which are then used to release the orders in the order pool. They show
that their approach outperforms static order release mechanisms by yielding lower costs, lateness
and standard deviation of lateness, but conclude that research using average reward reinforcement
learning methods harbor optimisation potential for the order release problem.
%
And finally, \citet{schneckenreither2020order} present a flow time estimation procedure to set lead
times dynamically using an artificial neural network, which is used to forecast flow times. By
implementing a rolling horizon order release simulation of the proceeding periods they lift their
approach to a predictive lead time management approach, which is able to detect backorders of future
periods and reacts by releasing orders earlier. Nonetheless, their method is unable to foresee the
triggering of the lead time syndrome and therefore they introduce an upper lead time bound to
prevent the negative effects of the lead time syndrome. Their model outperforms static and reactive
lead time management approaches, especially for scenarios with high utilisation and high variability
in processing times.

As reinforcement learning stems from dynamic programming the future system state is by design
considered as a main driver of decision making in the current period.
% Reinforcement learning is an optimisation technique that stems from dynamic programming and
Its goal is to find the best stationary policy for a given problem. % This policy is usually provided
% by assessing current and future states (or state-action pairs) of an underlying Markov Decision
% Process (MDP).
The advantage of reinforcement learning over dynamic programming is that (i) the problem
space is explored by an agent and thus only expectantly interesting parts of the problem space need
to be assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary
as the states are evaluated by consecutively observed states solely.

@@

Therefore, most applications that incorporate and directly reflect costs or profit in the reward
function use average reward reinforcement learning. \citet{aydin2000dynamic} use it in the setting
of scheduling, while in a series of papers Mahadevan et al. investigated several problem domains
starting with simple MDPs
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
After these foundational works for average reward reinforcement learning they introduced a
continuous time average reward reinforcement learning algorithm named SMART
\citep{Mahadevan97_SelfimprovingFactorySimulationUsingContinuoustimeAveragerewardReinforcementLearning}.
Applications of SMART reach from the optimisation of queuing systems
\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies,Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning},
maintenance of an inventory system
\citep{Das99_SolvingSemiMarkovDecisionProblemsUsingAverageRewardReinforcementLearning} to optimising
transfer line in terms of maximizing demand, while keeping inventory levels of unfinished product as
low as possible \citep{Mahadevan98_OptimizingProductionManufacturingUsingReinforcementLearning}.
However, in practise usually decision have to be made on a daily basis
\citep{enns2004work,GeldersvanW1982}. Therefore, we refrain from this adaption and concentrate on
standard MDPs only. Furthermore, often continuous-time semi-MDP problems can be converted through
uniformisation into equivalent discrete time instances
\citep[see][]{Puterman94,bertsekas1995dynamic}.
